Interpretable Machine Learning for Science \\ with PySR and SymbolicRegression.jl

### 2023-05-02 src/tex/ms.tex
Get minted to finalize a cache
Add to benchmark discussion
Grammar fixes

\usepackage{minted}
\setminted[]{
  frame=lines,
  framesep=2mm,
  bgcolor=bg,
  baselinestretch=1.2,
  fontsize=\footnotesize,
  breaklines,
  breakafter=d,
  breakbefore=a,
}
\usemintedstyle{solarized-light}
\definecolor{bg}{rgb}{0.97, 0.97, 0.97}

@@

Additional techniques include~\cite{jinBayesianSymbolicRegression2020,guimeraBayesianMachineScientist2020} which propose a Markov chain Monte Carlo-like sampling for \sr;
\cite{petersenDeepSymbolicRegression2021,liNeuralguidedSymbolicRegression2019} which trains deep generative models to propose mutated expressions;
\cite{dascoliDeepSymbolicRegression2022,kamiennyEndtoendSymbolicRegression2022} which pre-train a large transformer model on billions of synthetic examples, in order to auto-regressively generate expressions from data;
\cite{udrescuAIFeynmanParetooptimal2020,liuAIPoincareMachine2020,wetzelDiscoveringSymmetryInvariants2020} which develop techniques for discovering symmetries with \sr;
\cite{sahooLearningEquationsExtrapolation2018} which optimizes a sparse set of coefficients in a very large pre-defined expression with stochastic gradient descent; and finally~\cite{atkinsonDatadrivenDiscoveryFreeform2019,rossBenchmarkingMachineLearning2022} which apply genetic algorithms to a flexible space of differential equations.

Apart from trees, a few other data structures are used for representing symbolic expressions, such as imperative representations (``Linear Genetic Programming'')~\citep[e.g.,][]{brameierComparisonLinearGenetic2001,guvenLinearGeneticProgramming2009,maEvolvingSymbolicDensity2022,diasDescribingQuantumInspiredLinear2012}, and sequence representations~\citep[e.g.,][]{liuAIPoincareMachine2020,dascoliDeepSymbolicRegression2022,kamiennyEndtoendSymbolicRegression2022}.
One of the most popular alternatives to genetic algorithms on trees, especially for the study of PDEs, is ``SINDy'' (Sparse Identification of Nonlinear Dynamics) ~\cite{bruntonDiscoveringGoverningEquations2016,rudyDatadrivenDiscoveryPartial2017,championDatadrivenDiscoveryCoordinates2019} which represents expressions as linear combination of a dictionary of fixed nonlinear expressions.
Similar ideas of a linear basis of terms is used in the initialization of the ``Operon'' genetic algorithm~\cite{burlacuOperonEfficientGenetic2020}, as well as in ``FFX''~\cite{mcconaghyFFXFastScalable2011}.
This strategy can also be combined with deep learning~\citep[e.g.,][]{luschDeepLearningUniversal2018,championDatadrivenDiscoveryCoordinates2019,bothDeepMoDDeepLearning2019,chenDeepLearningPhysical2020,rackauckasUniversalDifferentialEquations2020} or other techniques to find expressions on high-dimensional data~\citep[e.g.,][]{rossBenchmarkingMachineLearning2022}.
Additional techniques include mixed integer nonlinear programming~\citep{cozadGlobalMINLPApproach2018} and greedy tree searches~\citep{defrancaGreedySearchTree2018}, and combining theory as a type of constraint over the space of models~\cite{cornelioAIDescartesCombining2021}.

@@

\pysr is a multi-population evolutionary algorithm with multiple evolutions performed asynchronously.
The main loop of \pysr, operating on each population independently, is a classic evolutionary algorithm based on tournament selection for individual selection (first introduced in the thesis of~\cite{brindleGeneticAlgorithmsFunction1980} and generalized in~\cite{goldbergComparativeAnalysisSelection1991}),
and several different mutations and crossovers for generating new individuals.
Evolutionary algorithms for \sr was first popularized in~\cite{kozaGeneticProgrammingMeans1994}, and has dominated the development of flexible \sr algorithms since.
\pysr makes several modifications to this classic approach, some of which are motivated by results in recent work.

@@

A simple evolutionary algorithm proceeds as follows:
\begin{easylist}
\ListProperties(Margin1=1cm,Hang1=true,Indent1=.5cm,Numbers1=a,Start1=1,Align=0.5cm)
# Assume one has a population of individuals, a fitness function, and a set of mutation operators.
\label{step:assumptions}
# Randomly select an $n_s$-sized subset of individuals from the population (e.g., classical tournament selection uses $n_s=2$, but larger tournaments are also allowed).
\label{step:sampling}
# Run the tournament by evaluating the fitness of every individual in that subset.
\label{step:prep_tournament}
# Select the fittest individual as the winner with probability $p$. Otherwise, remove this individual from the subset and repeat this step again.
If there is one remaining, select it.
Thus the probability will roughly be $p, p(1-p), p(1-p)^2, ...$ for the first, second, third fittest, and so on. 
\label{step:select}
# Create a copy of this selected individual, and apply a randomly-selected mutation from a set of possible mutations.
\label{step:mutate}
# Replace a member of the population with the mutated individual. Here, one would replace the weakest of the population or subset.
\label{step:replace}
\end{easylist}

This algorithm is computationally efficient and allows for massive paralellization by splitting step~\ref{step:sampling} into groups of individuals which can evolve independently, with asynchronous ``migration'' between the groups.
For the first of the modifications, \we replace the eldest member of a population in step~\ref{step:mutate}, rather than the individual with the lowest fitness, which is known as ``age-regularized'' evolution~\cite{realRegularizedEvolutionImage2019}.
We record the age from the Unix time at which the individual was created during step~\ref{step:mutate}.
Its inclusion in \pysr was motivated by the impressive Neural Architecture Search results of~\cite{realRegularizedEvolutionImage2019,realAutoMLZeroEvolvingMachine2020a} --- though it is worth noting similar regularizations used earlier in SR, e.g.,~\cite{hornbyALPSAgelayeredPopulation2006} and~\cite{schmidtAgefitnessParetoOptimization2010} --- which found this simple modification could prevent early convergence without hurting performance.
This prevents the population from specializing too quickly and getting stuck in a local minima of the search space.
The implementation in~\cite{realAutoMLZeroEvolvingMachine2020a} uses $p=1$ and explores $4\leq n_s \leq 64$, though here \we allow $p$ to be less than $1$, among several other differences.

@@

\We emphasize that the stability of an \sr algorithm and software implementation itself has central importance to practical use in science, and therefore, unstable runs which either produce an undefined expression, or crash, are \textit{included} in the results dataset and the primary score.
However, one should also note that because \we are the authors of \pysr, we are therefore more likely to be running in a stable environment than, say, a package which \we are unfamiliar with.
For example, as seen in \cref{tab:results}, DSR sees improved results if one ignores the failed runs, although there were still some expressions it struggled on.
Thus, detailed itemizations of these results are given in \cref{tab:results}, to allow for a more nuanced understanding of the results.

>>>>

\usepackage[finalizecache]{minted}
\setminted[]{
  frame=lines,
  framesep=2mm,
  bgcolor=bg,
  baselinestretch=1.2,
  fontsize=\footnotesize,
  breaklines,
  breakafter=d,
  breakbefore=a,
}
\usemintedstyle{solarized-light}
\definecolor{bg}{rgb}{0.97, 0.97, 0.97}

@@

Additional techniques include~\cite{jinBayesianSymbolicRegression2020,guimeraBayesianMachineScientist2020} which propose a Markov chain Monte Carlo-like sampling for \sr;
\cite{petersenDeepSymbolicRegression2021,liNeuralguidedSymbolicRegression2019} which trains deep generative models to propose mutated expressions;
\cite{dascoliDeepSymbolicRegression2022,kamiennyEndtoendSymbolicRegression2022} which pre-train a large transformer model on billions of synthetic examples, in order to autoregressively generate expressions from data;
\cite{udrescuAIFeynmanParetooptimal2020,liuAIPoincareMachine2020,wetzelDiscoveringSymmetryInvariants2020} which develop techniques for discovering symmetries with \sr;
\cite{sahooLearningEquationsExtrapolation2018} which optimizes a sparse set of coefficients in a very large pre-defined expression with stochastic gradient descent; and finally~\cite{atkinsonDatadrivenDiscoveryFreeform2019,rossBenchmarkingMachineLearning2022} which apply genetic algorithms to a flexible space of differential equations.

Apart from trees, a few other data structures are used for representing symbolic expressions, such as imperative representations (``Linear Genetic Programming'')~\citep[e.g.,][]{brameierComparisonLinearGenetic2001,guvenLinearGeneticProgramming2009,maEvolvingSymbolicDensity2022,diasDescribingQuantumInspiredLinear2012}, and sequence representations~\citep[e.g.,][]{liuAIPoincareMachine2020,dascoliDeepSymbolicRegression2022,kamiennyEndtoendSymbolicRegression2022}.
One of the most popular alternatives to genetic algorithms on trees, especially for the study of PDEs, is ``SINDy'' (Sparse Identification of Nonlinear Dynamics) ~\cite{bruntonDiscoveringGoverningEquations2016,rudyDatadrivenDiscoveryPartial2017,championDatadrivenDiscoveryCoordinates2019} which represents expressions as linear combination of a dictionary of fixed nonlinear expressions.
Similar ideas of a linear basis of terms is used in the initialization of the ``Operon'' genetic algorithm~\cite{burlacuOperonEfficientGenetic2020}, as well as in ``FFX''~\cite{mcconaghyFFXFastScalable2011}.
This strategy can also be combined with deep learning~\citep[e.g., ][]{luschDeepLearningUniversal2018,championDatadrivenDiscoveryCoordinates2019,bothDeepMoDDeepLearning2019,chenDeepLearningPhysical2020,rackauckasUniversalDifferentialEquations2020} or other techniques to find expressions on high-dimensional data~\citep[e.g., ][]{rossBenchmarkingMachineLearning2022}.
Additional techniques include mixed integer nonlinear programming~\citep{cozadGlobalMINLPApproach2018} and greedy tree searches~\citep{defrancaGreedySearchTree2018}, and combining theory as a type of constraint over the space of models~\cite{cornelioAIDescartesCombining2021}.

@@

\pysr is a multi-population evolutionary algorithm with multiple evolutions performed asynchronously.
The main loop of \pysr, operating on each population independently, is a classic evolutionary algorithm based on tournament selection for individual selection (first introduced in the thesis of~\cite{brindleGeneticAlgorithmsFunction1980} and generalized in~\cite{goldbergComparativeAnalysisSelection1991}),
and several mutations and crossovers for generating new individuals.
Evolutionary algorithms for \sr was first popularized in~\cite{kozaGeneticProgrammingMeans1994}, and has dominated the development of flexible \sr algorithms since.
\pysr makes several modifications to this classic approach, some of which are motivated by results in recent work.

@@

A simple evolutionary algorithm proceeds as follows:
\begin{easylist}
\ListProperties(Margin1=1cm,Hang1=true,Indent1=.5cm,Numbers1=a,Start1=1,Align=0.5cm)
# Assume one has a population of individuals, a fitness function, and a set of mutation operators.
\label{step:assumptions}
# Randomly select an $n_s$-sized subset of individuals from the population (e.g., classical tournament selection uses $n_s=2$, but larger tournaments are also allowed).
\label{step:sampling}
# Run the tournament by evaluating the fitness of every individual in that subset.
\label{step:prep_tournament}
# Select the fittest individual as the winner with probability $p$. Otherwise, remove this individual from the subset and repeat this step again.
If there is one remaining, select it.
Thus, the probability will roughly be $p, p(1-p), p(1-p)^2, \ldots$ for the first, second, third fittest, and so on. 
\label{step:select}
# Create a copy of this selected individual, and apply a randomly-selected mutation from a set of possible mutations.
\label{step:mutate}
# Replace a member of the population with the mutated individual. Here, one would replace the weakest of the population or subset.
\label{step:replace}
\end{easylist}

This algorithm is computationally efficient and allows for massive parallelization by splitting step~\ref{step:sampling} into groups of individuals which can evolve independently, with asynchronous ``migration'' between the groups.
For the first of the modifications, \we replace the eldest member of a population in step~\ref{step:mutate}, rather than the individual with the lowest fitness, which is known as ``age-regularized'' evolution~\cite{realRegularizedEvolutionImage2019}.
We record the age from the Unix time at which the individual was created during step~\ref{step:mutate}.
Its inclusion in \pysr was motivated by the impressive Neural Architecture Search results of~\cite{realRegularizedEvolutionImage2019,realAutoMLZeroEvolvingMachine2020a} --- though it is worth noting similar regularizations used earlier in SR, e.g.,~\cite{hornbyALPSAgelayeredPopulation2006} and~\cite{schmidtAgefitnessParetoOptimization2010} --- which found this simple modification could prevent early convergence without hurting performance.
This prevents the population from specializing too quickly and getting stuck in a local minimum of the search space.
The implementation in~\cite{realAutoMLZeroEvolvingMachine2020a} uses $p=1$ and explores $4\leq n_s \leq 64$, though here \we allow $p$ to be less than $1$, among several other differences.

@@

\We emphasize that the stability of an \sr algorithm and software implementation itself has central importance to practical use in science, and therefore, unstable runs which either produce an undefined expression, or crash, are \textit{included} in the results dataset and the primary score.
However, one should also note that because \we are the authors of \pysr, we are therefore more likely to be running in a stable environment than, say, a package which \we are unfamiliar with.
For example, as seen in \cref{tab:results}, DSR sees improved results if one ignores the failed runs, although there were still some expressions it struggled on.
Thus, detailed itemizations of these results are given in \cref{tab:results}, to allow for a more nuanced understanding of the results.
Also note that this benchmark does not explicitly measure other metrics which are considered in other benchmarks, such as number of evaluations (where, e.g., genetic algorithms are very inefficient), nor does it consider the numerical accuracy of the expressions.
